---
  title: "Genotype Information Extraction"
  author: "Gepoliano Chaves"
  date: "June 24th, 2024"
  output:
    html_document:
      df_print: paged
      toc: true
  bibliography: references.bib
---

\

## Summary

\

In this activity we extract genetic information from VCF files to generate genotype file that will be used in the machine learning framework.

\

## Load libraries

\

```{r, message=F, warning=F}
library(maditr)
library(kableExtra)
library(dplyr)
library(stringr)
library(tibble)
library(reshape2)
```

\

## 1. Extracting Genotypes


\

https://stackoverflow.com/questions/7743898/how-to-use-cast-on-a-data-frame

* Change from the long format to the wide format:

https://www.statology.org/cast-in-r/

https://dplyr.tidyverse.org/reference/relocate.html

\

The file3.txt below was generated using the query parameter in the bcftools used in the previous notebook. This file needs to be filled until I figure out what is wrong with that filling. Genotypes are 0 and 1 for file4.txt. This chunk was being evaluated before I started the PRS analysis

```{r read file, eval=T}
mt_extracted <- read.table("data/file4.txt", sep = "\t", header = T)

names(mt_extracted)[1]<-paste("sample")
names(mt_extracted)[2]<-paste("chr")
names(mt_extracted)[3]<-paste("pos") 
names(mt_extracted)[4]<-paste("genotype")

dim(mt_extracted)
```

```{r read file no chr, eval=F}
mt_extracted <- read.table("data/genotypes.txt", sep = " ", header = F)

names(mt_extracted)[1]<-paste("pos")
names(mt_extracted)[2]<-paste("genotype")
names(mt_extracted)[3]<-paste("sample") 

dim(mt_extracted)
```

\

Fix empty spaces in object mt_extracted 

```{r, eval=T}
# Split Each Sample
split_sample_name_df <- str_split_fixed(mt_extracted$sample, " ", 2)
# Construct Vector Maintaining parts of interest
split_vector_sample_names <- paste0(split_sample_name_df[, 1])
# Add vector to DF, after column of interest
mt_extracted <- add_column(mt_extracted, sampleID = split_vector_sample_names, .after = "sample")
mt_extracted <- mt_extracted %>% dplyr::select(-c(sample))
names(mt_extracted)[names(mt_extracted) == 'sampleID'] <- 'sample'
head(mt_extracted)
```

Fix empty spaces in object mt_extracted 

```{r, eval=F}
# Split Each Sample
split_sample_name_df <- str_split_fixed(mt_extracted$sample, "_", 2)
# Construct Vector Maintaining parts of interest
split_vector_sample_names <- paste0(split_sample_name_df[, 1])
# Add vector to DF, after column of interest
mt_extracted <- add_column(mt_extracted, sampleID = split_vector_sample_names, .after = "sample")
mt_extracted <- mt_extracted %>% dplyr::select(-c(sample))
names(mt_extracted)[names(mt_extracted) == 'sampleID'] <- 'sample'
head(mt_extracted)
```

\

### Selected columns included

\

Select only the columns that will make the dataframe compatible with the dcast function of the reshape2 library.
Reshape package has functions dcast and melt. In this part, we change from the long format to the wide format using the dcast() function in R. This chunk was being evaluated before I started the PRS analysis

```{r select columns for function dcast(), eval=T}
library(maditr)
mt_selected <- mt_extracted %>% dplyr::select(sample, pos, genotype)

library(reshape2)

mt_dcast <- dcast(mt_selected, sample ~ pos)

head(mt_dcast)
dim(mt_dcast)
```

\

Use function dcast

```{r, eval=F}
library(maditr)
mt_selected <- mt_extracted %>% dplyr::select(sample, pos, genotype)
library(reshape2)
mt_dcast_gt <- dcast(mt_selected, sample ~ pos)
```



\

This chunk is from notebook 03-A-merge_vcfs.Rmd, that was at first with this notebook. Chunk "Merge sampleIDs and mt_dcast", below, needs object sampleIDs, that is being loaded here. 

```{r sample IDs from 03-A, eval=T}
sampleIDs <- read.table("data/sampleIDs.txt")
names(sampleIDs)[1]<-paste("GenBankID")
names(sampleIDs)[2]<-paste("sampleName")
head(sampleIDs)
```

\

```{r sample IDs with 41 samples, eval=F}
sampleIDs <- read.table("data/sampleIDs.txt")
names(sampleIDs)[1]<-paste("GenBankID")
names(sampleIDs)[2]<-paste("sampleName")
head(sampleIDs)
```

\

Order sample ID columns so check columns are equal is possible.

```{r}
mt_dcast <- mt_dcast[order(mt_dcast$sample),]
sampleIDs <- sampleIDs[order(sampleIDs$sampleName),]
```

\

```{r, eval=F}
mt_dcast_gt <- mt_dcast_gt[order(mt_dcast_gt$sample),]
sampleIDs <- sampleIDs[order(sampleIDs$GenBankID),]
```

\

Or can also use identical:

```{r}
identical(mt_dcast[['sample']],sampleIDs[['sampleName']])
```

\

```{r, eval=F}
identical(mt_dcast_gt[['sample']],sampleIDs[['GenBankID']])
```

\

Rename samples with GenBank IDs. This chunk was being evaluated before I started the PRS analysis

```{r Merge sampleIDs and mt_dcast , eval=T, message=F, warning=F}

library(data.table)

dt1         <- data.table(mt_dcast , key = "sample") 
dt2         <- data.table(sampleIDs, key = "sampleName")
joined_df_1 <- dt1[dt2]
mt_dcast    <- joined_df_1
mt_dcast <- mt_dcast %>% dplyr::relocate(GenBankID, .before = sample)
mt_dcast <- mt_dcast %>% dplyr::select(-c(sample))
head(mt_dcast)
```

\

```{r, eval=T, message=F, warning=F, eval=F}

library(data.table)

dt1         <- data.table(mt_dcast , key = "sampleName") 
dt2         <- data.table(sampleIDs, key = "sampleName")
joined_df_1 <- dt1[dt2]
mt_dcast    <- joined_df_1
mt_dcast <- mt_dcast %>% dplyr::relocate(sample, .before = sample)
mt_dcast <- mt_dcast %>% dplyr::select(-c(sample))
head(mt_dcast)
```

\

this chunk needs to call the metadataIDs object:

```{r sample IDs metadataIDs.txt, eval=T}
metadataIDs <- read.table("data/metadataIDs.txt")
names(metadataIDs)[1]<-paste("GenBankID")
names(metadataIDs)[2]<-paste("sampleName")
names(metadataIDs)[3]<-paste("origin")
```

\

In this part we edit column names to include the format MT_XXXXXXX_BP. This chunk was being evaluated before I started the PRS analysis

```{r, eval=T}
colnames(mt_dcast) <- paste("MT", colnames(mt_dcast), "BP", sep= "_")
names(mt_dcast)[1]<-paste("GenBankID")
head(mt_dcast)
```


\

Insert column with labels. This chunk was being evaluated before I started the PRS analysis

```{r, eval=T}
head(mt_dcast)
```

\

Include the origin column

```{r, eval=T}
mt_dcast <- mt_dcast %>% dplyr::mutate(origin = ifelse(row_number() < 10, "America", "Africa" ))
```

\

Check insertion of the origin column

```{r, eval=T}
head(mt_dcast)
```

\

Relocate the column that indicated the origin. This chunk was being evaluated before I started the PRS analysis

```{r, eval=T}
mt_dcast <- mt_dcast %>% dplyr::relocate(origin, .after = GenBankID)
```

\

Check relocation of the origin column

```{r, eval=T}
head(mt_dcast)
```

\

## 2. Changing Genotypes to Binary

\

To change genotypes to 1/0, I used this @stackOverflow2021 reference.

This chunk was being evaluated before I started the PRS analysis

```{r, eval=T}
mt_dcast

mt_dcast <- mt_dcast %>% mutate_all(funs(replace(., .== "C>T", "1")))%>% 
  mutate_all(funs(replace(., .== "T>C", "1")))%>% 
  mutate_all(funs(replace(., .== "C>A", "1")))%>% 
  mutate_all(funs(replace(., .== "G>A", "1")))%>% 
  mutate_all(funs(replace(., .== "A>G", "1")))%>% 
  mutate_all(funs(replace(., .== "A>T", "1")))

mt_dcast
```

\

### 2.1. Visualize Table

\

Take a look at table with genotypes changed

This chunk was being evaluated before I started the PRS analysis

```{r look at new genotype table, eval=T}

kable(mt_dcast, caption="New Genotype Table (Reference > Alteration)") %>%
  kable_styling("striped", full_width = F, font_size = 12) %>%
  scroll_box(width = "100%", height = "600px")

```

Check dimensions of dataframe

\

This chunk was being evaluated before I started the PRS analysis

```{r, eval=F}
dim(mt_dcast)
```


Write to file

This chunk was being evaluated before I started the PRS analysis

```{r save rds, eval=F}
saveRDS(mt_dcast, file = "./data/mt_dcast.rds")
```

Save txt file. Do not leave this chunk on so it can not randomly modify other notebooks that use the mt_dcast object.

```{r save txt, eval=F}
write.table(mt_dcast,
            file = "./data/mt_dcast.txt",
            sep = "\t",
            row=F,
            quote = F,
            col.names = T)
```

\

### 2.2. Process and save genotype file

\

This chunk was being evaluated before I started the PRS analysis

```{r, eval=F}
mt_dcast_appended_binary <- mt_dcast_appended %>% mutate_all(funs(replace(., .== "C=C", "0")))%>% 
  mutate_all(funs(replace(., .== "C>T", "1")))%>% 
  mutate_all(funs(replace(., .== "T=T", "0")))%>% 
  mutate_all(funs(replace(., .== "T>C", "1")))%>% 
  mutate_all(funs(replace(., .== "G=G", "0")))%>% 
  mutate_all(funs(replace(., .== "G>A", "1")))%>%
  mutate_all(funs(replace(., .== "A=A", "0")))%>%
  mutate_all(funs(replace(., .== "A>G", "1")))
```

Take a look at mt_dcast_appended_binary table

\

This chunk was being evaluated before I started the PRS analysis

```{r look at table, eval=F}
library(kableExtra)

kable(mt_dcast_appended_binary, caption="Genotype Table (Reference > Alteration)") %>%
  kable_styling("striped", full_width = F, font_size = 12) %>%
  scroll_box(width = "100%", height = "600px")
```

Save mt_dcast_appended_binary

This chunk was being evaluated before I started the PRS analysis

```{r save mt_dcast_appended_binary, eval=F}
write.table(mt_dcast_appended_binary,
            file = "./data/mt_dcast_appended_binary.txt",
            sep = "\t",
            row=F,
            quote = F,
            col.names = T)
```


\

## 3. References

\

<div id="refs"></div>

\

## 4. Session Info

\

```{r session}
sessionInfo()
```