Load library

```{r}
library("msa")
```
\

Read fasta file

```{r}
mt_brazilian_capital <- readDNAStringSet(file="./data/mt_brazilian_capital.fasta")
```

\

Look at the mt_brazilian_capital object

```{r}
mt_brazilian_capital
```

\

Align sequences

```{r}
alignment_mt_brazilian_capital <- msa(mt_brazilian_capital)
```

\

Visualize alignment object

```{r}
alignment_mt_brazilian_capital
```

\

### Distance Matrix and Phylogenetic Tree of Partial Sequence

\

This is necessary so function dist.ml can take the DNA.bin object and process the Distance Matrix. Function dist.ml() is from package phangorn

```{r}
library("phangorn")
alignment_mt_brazilian_capital_dist.ml <- dist.ml(as.DNAbin(alignment_mt_brazilian_capital))
```

\

### Plot distance matrix

\
Function table.paint() is from package adegenet. Load library adegenet:

```{r}
library("adegenet")
dm_alignment_mt_brazilian_capital_dist.ml_df <- as.data.frame(as.matrix(alignment_mt_brazilian_capital_dist.ml))
table.paint(dm_alignment_mt_brazilian_capital_dist.ml_df, cleg=0, clabel.row=.5, clabel.col=.5)
```


### Construct Phylogenetic Trees

\

Phylogenetic tree UPGMA method

```{r}
treeUPGMA_dm_alignment_mt_brazilian_capital_dist.ml_df <- upgma(dm_alignment_mt_brazilian_capital_dist.ml_df)
```

\

### Plot Phylogenetic Tree mtDNA UPGMA

```{r}
plot(treeUPGMA_dm_alignment_mt_brazilian_capital_dist.ml_df, main="UPGMA", col="red")
```
\


### Save the plot

```{r}
library(ggplot2)
p_treeUPGMA <- plot(treeUPGMA_dm_alignment_mt_brazilian_capital_dist.ml_df, main="UPGMA", col="red")

# Dimensions default to pixels for png()
png("./figures/treeUPGMA_plot.png", width = 2000, height = 2000, res=100)
plot(treeUPGMA_dm_alignment_mt_brazilian_capital_dist.ml_df, main="UPGMA", col="red")
dev.off()

```
\

### Visualize alignment




```{r}
library(ggmsa)
ggmsa(mt_brazilian_capital, 
      start = 40, end = 70, 
      char_width = 0.5, 
      seq_name = T) + geom_seqlogo() + geom_msaBar()
```

\

Save the plot

```{r}
# Dimensions default to pixels for png()
png("./figures/alignment_plot.png", width = 2000, height = 2000, res=150)
ggmsa(mt_brazilian_capital, 
      start = 40, end = 70, 
      char_width = 0.5, 
      seq_name = T) + geom_seqlogo() + geom_msaBar()
dev.off()
```





