
### 4. Example datasets

\

```{r, warning=F, message=F}
library("ggenealogy")
library("dplyr")
library("readxl")
```

\

```{r}
data("sbGeneal", package = "ggenealogy")
data("statGeneal", package = "ggenealogy")
statGenealEP <- statGeneal %>% filter(parent != "")
statIG <- dfToIG(statGenealEP)
uCP <- na.omit(c(statGeneal$child, statGeneal$parent))
uChild <- unique(na.omit(statGeneal$child))
uParent <- unique(na.omit(statGeneal$parent))
```

\

Specify sheet with a number or name

```{r}
brGeneal <- read_excel("../geraldo/data/brGeneal.xlsx", sheet = "Sheet3")
```


\

### 6. Generating a graphical object

```{r}
sbIG <- dfToIG(sbGeneal)
getBasicStatistics(sbIG)
```

\

```{r}
brIG <- dfToIG(brGeneal)
```


\

### 7. Plotting a shortest path

\

```{r}
pathTN <- getPath("Tokyo", "Narow", sbIG, sbGeneal, "devYear")
plotPath(pathTN, sbGeneal, "devYear")
```

\

```{r}
pathBZ <- getPath("Bedford", "Zane", sbIG, sbGeneal, "devYear")

plotPath(pathBZ, sbGeneal, "devYear", fontFace = 2) +
  ggplot2::xlab("Development Year")
```

\

Descendents of the kings of Portugal from the first king, Afonso Henriques (Afonso I of Portugal), to Martim Afonso Chichorro, Governor of Chaves in Northern Portugual. We can use the official information of nobles in Portugal and Brazil to trace the ancestry of common Portuguese and Brazilian citizens and their families:

```{r}
pathTN_BR <- getPath("Afonso I de Portugal",
                     #"Capitão André Rodrigues Chaves",
                     #"Francisco Dias d'Ávila",
                     "Martim Afonso Chichorro",
                     brIG, brGeneal, "devYear")
plotPath(pathTN_BR, brGeneal, "devYear")
```



Now we plot the branch of Chaves immigrants to Brazil. 

\

Below we see immigration from Chaves in Portugal to Brazil that originated families and the City of Dias D'Ávila in Bahia and Babuí (Minas Gerais):

\

```{r}
pathTN_BR <- getPath("Garcia D'Ávila", "Maria Rita Chaves", 
                     brIG, brGeneal, "devYear")

plotPath(pathTN_BR, sbGeneal, "devYear", fontFace = 2) + ggplot2::xlab("Year of Birth")
```

\

```{r}
sbFilt <- sbGeneal[complete.cases(sbGeneal[1:3]), ]
sbFiltIG <- dfToIG(sbFilt)
pathCL <- getPath("Clark", "Lawrence", sbFiltIG, sbFilt, "yield")
plotPath(pathCL, sbFilt, "yield") + ggplot2::xlab("Yield")
pathCL2 <- getPath("Clark", "Lawrence", sbFiltIG, sbFilt, "devYear")
plotPath(pathCL2, sbFilt, "devYear", "yield") +
  ggplot2::xlab("Development Year") + ggplot2::ylab("Yield")
```

\

### 8. Superimposing a shortest path on a tree

\

```{r}
plotPathOnAll(pathTN, sbGeneal, sbIG, "devYear", bin = 3, 
              pathEdgeCol = "red", 
              nodeSize = 2.5, 
              pathNodeSize = 4) +
  ggplot2::theme(axis.text = ggplot2::element_text(size = 12), 
                 axis.title = ggplot2::element_text(size = 12))
```

\

```{r}
plotPathOnAll(pathTN_BR, brGeneal, brIG, "devYear", 
              pathEdgeCol = "red", 
              nodeSize = 2.5, 
              pathNodeSize = 4) +
  ggplot2::theme(axis.text = ggplot2::element_text(size = 12), 
                 axis.title = ggplot2::element_text(size = 12))
```

\

```{r}
plotPathOnAll(pathTN, sbGeneal, sbIG, "devYear", bin = 6,
              pathEdgeCol = "seagreen2", 
              nodeSize = 1, pathNodeSize = 3) + 
  ggplot2::xlab("Development Year")
```

\

```{r}
plotPathOnAll(pathCL, sbFilt, sbFiltIG, "yield", bin = 3, pathEdgeCol = "purple") + ggplot2::xlab("Yield")

plotPathOnAll(pathCL, sbFilt, sbFiltIG, "yield", "devYear", pathEdgeCol = "orange") + ggplot2::xlab("Yield") + ggplot2::ylab("Development Year")
```

\

### 9. Plotting ancestors and descendants by generation

\


Starting with Gonçalo Vaz Coutinho

\

```{r}
brGeneal <- read_excel("../geraldo/data/brGeneal.xlsx", sheet = "Sheet9")
plotAncDes("Pero do Campo Tourinho I", brGeneal, mAnc = 3, mDes = 7, vCol = "blue")
```

\

Starting with Estêvão Gil Tourinho

\

```{r}
brGeneal <- read_excel("../geraldo/data/brGeneal.xlsx", sheet = "Sheet9")
plotAncDes("Estêvão Gil Tourinho", brGeneal, mAnc = 1, mDes = 7, vCol = "blue")
```

\

Starting with Estêvão Gil Tourinho

\

```{r}
brGeneal <- read_excel("../geraldo/data/brGeneal.xlsx", sheet = "Sheet9")
plotAncDes("Gaspar Tourinho (Ilhéus)", brGeneal, mAnc = 0, mDes = 7, vCol = "blue")
```

\

Starting with Cosme de Sá Tourinho

\

```{r}
brGeneal <- read_excel("../geraldo/data/brGeneal.xlsx", sheet = "Sheet9")
plotAncDes("Cosme de Sá Tourinho", brGeneal, mAnc = 0, mDes = 7, vCol = "blue")
```

\

Starting with Maria Tereza da Paz

\

```{r}
brGeneal <- read_excel("../geraldo/data/brGeneal.xlsx", sheet = "Sheet9")
plotAncDes("Cosme de Sá Tourinho", brGeneal, mAnc = 0, mDes = 7, vCol = "blue")
```

\

Starting with Paula Camacho

\

```{r}
brGeneal <- read_excel("../geraldo/data/brGeneal.xlsx", sheet = "Sheet9")
plotAncDes("Paula Camacho", brGeneal, mAnc = 0, mDes = 7, vCol = "blue")
```



\

Starting with Fernão Dias Paes Leme (Filho)

\

```{r}
brGeneal <- read_excel("../geraldo/data/brGeneal.xlsx", sheet = "Sheet9")
plotAncDes("Fernão Dias Paes Leme (Filho)", brGeneal, mAnc = 7, mDes = 7, vCol = "blue")
```

\

Starting with João Tourinho Maciel

\

```{r}
brGeneal <- read_excel("../geraldo/data/brGeneal.xlsx", sheet = "Sheet9")
plotAncDes("João Tourinho Maciel", brGeneal, mAnc = 7, mDes = 7, vCol = "blue")
```


\


Starting with Pedro Dias Paes Leme

\

```{r}
brGeneal <- read_excel("../geraldo/data/brGeneal.xlsx", sheet = "Sheet9")
plotAncDes("Pedro Dias Paes Leme", brGeneal, mAnc = 7, mDes = 7, vCol = "blue")
```

\

Starting with Capitão-Mor Fernão Dias Paes Leme (O Caçador de Esmeraldas)

\

```{r}
brGeneal <- read_excel("../geraldo/data/brGeneal.xlsx", sheet = "Sheet9")
plotAncDes("Capitão-Mor Fernão Dias Paes Leme (O Caçador de Esmeraldas)", brGeneal, mAnc = 7, mDes = 7, vCol = "blue")
```

\

Starting with Capitão-Mor Fernão Dias Paes Leme (O Caçador de Esmeraldas) with 3 generations ascending:

\

```{r}
brGeneal <- read_excel("../geraldo/data/brGeneal.xlsx", sheet = "Sheet9")
plotAncDes("Capitão-Mor Fernão Dias Paes Leme (O Caçador de Esmeraldas)", brGeneal, mAnc = 1, mDes = 7, vCol = "blue")
```

\

Starting with Pedro Dias Paes Leme:

\

```{r}
brGeneal <- read_excel("../geraldo/data/brGeneal.xlsx", sheet = "Sheet9")
plotAncDes("Pedro Dias Paes Leme", brGeneal, mAnc = 1, mDes = 7, vCol = "blue")
```

\

Starting with Izabel Paes da Silva Irmã:

\

```{r}
brGeneal <- read_excel("../geraldo/data/brGeneal.xlsx", sheet = "Sheet9")
plotAncDes("Izabel Paes da Silva Irmã", brGeneal, mAnc = 0, mDes = 7, vCol = "blue")
```

\

Starting with Alice Rocha:

\

```{r}
brGeneal <- read_excel("../geraldo/data/brGeneal.xlsx", sheet = "Sheet9")
plotAncDes("Alice Rocha", brGeneal, mAnc = 0, mDes = 7, vCol = "blue")
```


