---
  title: "Confusion Table"
  author: "Gepoliano Chaves"
  date: "June 25th, 2024"
  output:
    html_document:
      df_print: paged
      toc: true
  bibliography: references.bib
---

\

# Summary 

\

This notebook was generated following documentation provideded by @Hutson2024.

\

## Running Time

\

```{r}
start_time <- Sys.time()
```


### Preparing the ML model to then evaluate

```{r, warning=F, message=F}
library(caret)
library(dplyr)
library(mlbench)
library(tidyr)
library(e1071)
library(randomForest)

# Load in the iris data set for this problem 
data(iris)
df <- iris
# View the class distribution, as this is a multiclass problem, we can use the multi-uclassification data table builder
table(iris$Species)

```

Splitting the data into train and test splits:

```{r}
train_split_idx <- caret::createDataPartition(df$Species, p = 0.75, list = FALSE)
# Here we define a split index and we are now going to use a multiclass ML model to fit the data
train <- df[train_split_idx, ]
test <- df[-train_split_idx, ]
str(train)

```

This now creates a 75% training set for training the ML model and we are going to use the remaining 25% as validation data to test the model.

```{r}
rf_model <- caret::train(Species ~ .,
                         data = df,
                         method = "rf",
                         metric = "Accuracy")

rf_model
```

The model is relatively accurate. This is not a lesson on machine learning, however we now know how well the model performs on the training set, we need to validate this with a confusion matrix. The Random Forest shows that it has been trained on greater than >2 classes so this moves from a binary model to a multi-classification model. The functions contained in the package work with binary and multiclassification methods.

### Using the native Confusion Matrix in CARET

* The native confusion matrix is excellent in CARET, however it is stored as a series of list items that need to be utilised together to compare model fit performance over time to make sure there is no underlying feature slippage and regression in performance. This is where my solution comes in.

* rf_class is a factor created using the predict() function
    + predict() takes on the **model**, the **newdata** and **type** parameters

* **preditions** is a data-frame constructed with the two columns used for the confusionMatrix() function.

```{r}
rf_class <- predict(rf_model, newdata = test, type = "raw") 
predictions <- cbind(data.frame(train_preds=rf_class, 
                                test$Species))
# Create a confusion matrix object
cm <- caret::confusionMatrix(predictions$train_preds, predictions$test.Species)
cm
print(cm)
```

### Using ConfusionTableR to collapse this data into a data frame

```{r}
# Implementing function to collapse data
library(ConfusionTableR)
mc_df <- ConfusionTableR::multi_class_cm(predictions$train_preds, predictions$test.Species,
                                         mode="everything")
# Access the reduced data for storage in databases
print(mc_df$record_level_cm)

glimpse(mc_df$record_level_cm)

```

This stores a list item. Here you can retrieve:

the confusion matrix, as this is generated automatically and does not require one to be fit beforehand, as in the previous example
the record_level_cm that can then be used to output data into a database
the confusion matrix numerical table
the datetime the list was created
To get the original confusion matrix:

```{r}
mc_df$confusion_matrix
```

To get the confusion matrix table:

```{r}
mc_df$cm_tbl
```

This data frame can now be used to store analyse these records over time i.e. looking at the machine learning statistics and if they depreciate or reduce upon different training runs.

### Using ConfusionTableR to collapse binary confusion matrix outputs

In this example we will use the breast cancer datasets, from mlbench to allow us to predict whether a new patient has cancer, dependent on the retrospective patterns in the data and the underlying data features.

#### Preparing data and fitting the model

```{r}
# Load in the data
library(dplyr)
library(ConfusionTableR)
library(caret)
library(tidyr)
library(mlbench)

# Load in the data
data("BreastCancer", package = "mlbench")
breast <- BreastCancer[complete.cases(BreastCancer), ] #Create a copy
breast <- breast[, -1]
breast$Class <- factor(breast$Class) # Create as factor
for(i in 1:9) {
 breast[, i] <- as.numeric(as.character(breast[, i]))
}
```

We now have our stranded patient model ready. Now we will fit a confusion matrix to this and use the tools in ConfusionTableR to output to a record level list, as observed in the previous section and to build a visualisation of the confusion matrix.

#### Predicting the class labels using the training dataset

This snippet shows how to achieve this:

```{r}
#Perform train / test split on the data
train_split_idx <- caret::createDataPartition(breast$Class, p = 0.75, list = FALSE)
train <- breast[train_split_idx, ]
test <- breast[-train_split_idx, ]
rf_fit <- caret::train(Class ~ ., data=train, method="rf")
#Make predictions to expose class labels
preds <- predict(rf_fit, newdata=test, type="raw")
predicted <- cbind(data.frame(class_preds=preds), test)
```

Now this is where we will use the package to visualise and reduce to a data frame.

### Binary Confusion Matrix Data Frame

The following example shows how this is implemented:

```{r}
bin_cm <- ConfusionTableR::binary_class_cm(predicted$class_preds, predicted$Class)
# Get the record level data
bin_cm$record_level_cm
glimpse(bin_cm$record_level_cm)
```

This is now in a data.frame class and can be used and saved as a single record to a database server to monitor confusion matrix performance over time.

### Visualising the confusion matrix

The last tool in the package produces a nice visual of the confusion matrix that can be used in presentations and papers to display the matrix and its associated summary statistics:

```{r}
ConfusionTableR::binary_visualiseR(train_labels = predicted$class_preds,
                                   truth_labels= predicted$Class,
                                   class_label1 = "Not Stranded", 
                                   class_label2 = "Stranded",
                                   quadrant_col1 = "#28ACB4", 
                                   quadrant_col2 = "#4397D2", 
                                   custom_title = "Breast Cancer Confusion Matrix", 
                                   text_col= "black")
```


# References

<div id="refs"></div>

# Session Info

```{r session}
sessionInfo()
```

```{r}
end_time <- Sys.time()
end_time - start_time
```

